---
title: "insect_biominformatics"
author: "Lia Domke"
date: "4/20/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script is based on the below tutorial by shawn wilkinson - some of the relevant information is brought over from that script, but it'd be good to review that as well. 
https://shaunpwilkinson.github.io/post/insect-vignette/

This script requires the ASV table from DADA2 tutorial
Ultimately you need to transform that ASV table to a DNAbin for classification. 
```{r}
#install.packages("insect")
library(insect)
library(tidyverse)
```

You need to download the classifier that is PRIMER SPECIFIC
This classifier is for MiFish

```{r}
# download.file("https://www.dropbox.com/s/fv3dpvws6zjvtib/classifier.rds?dl=1", 
#               destfile = "classifier.rds", mode = "wb")

# this downloads the classifier called classifier.rds into your working directory - only needs to be done once
```


First we need to classify sequences
The input for this is the seqtab.nochim
```{r}
# input read in seqtab.nochim
nochim <- read.csv("Data/eDNA_analysis/blast_DADA2_outputs/MiFish_ASV_seqtab_nochim.csv")
#nochim2 <- read.csv("csv_outputs/MiFish_ASV_seqtab_nochim_LD_04-20-23.csv")

asv <- read.csv("Data/eDNA_analysis/blast_DADA2_outputs/MiFish_ASVtable.csv")
```



```{r}
data(samoa) # the colum names should have the sequences
# samoa is the sample dataset used in the example tutorial
head(nochim)
# these two data types don't match need to adjust the nochim data to match how samoa is formated

# the row 1 is a seq related to ASV1 
# need to insert one row above the first named correctly

# okay lets extract the asv numbers by themselves 
asv.names <- nochim %>%
  data.frame() %>%
  mutate(sample.names = str_extract(X.ASV1, "ASV\\d*")) %>% # extract any instance of ASV and whatever followed
  dplyr::select(sample.names) %>%
  na.omit() %>%
  rbind("ASV1") %>%
  mutate(num = as.numeric(str_extract(sample.names, "\\d+"))) %>%
  mutate(chr = str_extract(sample.names, "ASV")) %>%
  arrange(num) %>%
  dplyr::select(sample.names)

nochim1.2 <- nochim %>%
  data.frame() %>%
  mutate(sample.names = str_extract(X.ASV1, "ASV\\d*")) %>%
  filter(is.na(sample.names)) %>%
  rename(seq = X.ASV1) %>%
  dplyr::select(seq)

# okay so now we've isolated the ASV and the seq and they've maintained order so just join back together

# now we need the sequences to be the column names
nochim.seq <- cbind(asv.names, nochim1.2) %>%
  pivot_wider(names_from = "seq", values_from = "sample.names")
```

```{r}
## get sequences from table column names
x <- char2dna(colnames(nochim.seq))
## name the sequences sequentially
names(x) <- paste0("ASV", seq_along(x))
## optionally remove column names that can flood the console when printed
colnames(asv) <- NULL 

## get sequences from table column names
x2 <- char2dna(colnames(nochim2))
## name the sequences sequentially
names(x2) <- paste0("ASV", seq_along(x2))
## optionally remove column names that can flood the console when printed
colnames(nochim2) <- NULL 
```


Next step is to load the classifier

this is an "insect" class object and is a large dendrogram with attributes for classifying sequences
```{r}
classifier <- readRDS("classifier.rds")
classifier
names(attributes(classifier))
```



Now we assign taxon IDS and CI to each ASV 

- classify function may take a minute or so
- right now ping = 1 which means it'll only return sequences with 100% identity are considered matches

**note** this can be run in parallel by setting: 
cores argument to 2 or more 
(setting cores = "autodetect") will have it run on one less than available. 

Other computational considerations:
- if choosing autodetect on a large classifier like COI 
then you MUST have 2 GB of available RAM per processor 

DADA2 output already has the sequence counts so we set tabulize = FALSE

In the case where a list of sequences containing replicates is to be processed, users can prefix the sequence names with their respective sample names, delimited with an underscore (e.g. “sample001_sequence001”) and set tabulize = TRUE. In this case, the classify function will automatically count and dereplicate the sequences, producing an output table with one column of sequence counts for each sample.


BELOW IS THE COMPUTATIONALLY INTENSIVE STEP MAKE SURE YOU'RE READY FOR IT:
```{r}
longDF <- classify(x, classifier, threshold = 0.8, tabulize = FALSE, species = "all", cores = "autodetect",
                   metadata = TRUE, ping = FALSE)

# try out using ping = True too
```


For DADA2 users, the ASV abundance table can now be transposed and appended to the table of taxonomic information if required:

```{r}
longDF <- cbind(longDF, t(nochim.seq))

#write.csv(longDF, "Data/insect_taxonomy_output.csv")
```

